---
title: "Worksheet 4a"
author: "Rica Marie Benliro"
date: "2023-11-01"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

1. The table below shows the data about shoe size and height. Create a data frame.

```{r}
shoe_size <- c(6.5,9.0,8.5,8.5,10.5,7.0,9.5,9.0,13.0,7.5,10.5,8.5,12.0,10.5,13.0,11.5,8.5,5.0,10.0,6.5,7.5,8.5,10.5,8.5,10.5,11.0,9.0,13.0)
height <- c(66.0,68.0,64.5,65.0,70.0,64.0,70.0,71.0,72.0,64.0,74.0,67.0,71.0,71.0,77.0,72.0,59.0,62.0,72.0,66.0,64.0,67.0,73.0,69.0,72.0,70.0,69.0,70.0)
gender <- c("F","F","F","F","M","F","F","F","M","F","M","F","M","M","M","M","F","F","M","F","F","M","M","F","M","M","M","M")

shoesizdf <- data.frame(shoe_size,height,gender)
shoesizdf
```

a. Describe the data.

```
#The "Household Data" table includes three main variables: "Shoe size" and "Height" (numerical in inches) and "Gender" (categorical). It consists of 32 observations, representing individuals in the household. While the data provides insights into shoe size and height within the household, it's important to note its limited scope to that specific household and the assumption of data quality. Despite this, it has potential applications in marketing research, clothing manufacturing, and genetic studies related to shoe size and height.
```


b. Create a subset by males and females with their corresponding shoe size and height. What its result? Show the R scripts.

```{r}
male <- subset(shoesizdf, gender == "M")
male
```
```{r}
female <- subset(shoesizdf, gender == "F")
female
```

c. Find the mean of shoe size and height of the respondents. Write the R scripts and its result.

```{r}
mean(shoesizdf$shoe_size)
```
```{r}
mean(shoesizdf$height)
```
d. Is there a relationship between shoe size and height? Why?
````
Yes. The data shows that people with heights of 70 and above have bigger show sizes compared to heights 69 and below.
```

2. Construct character vector months to a factor with factor() and assign the result to factor_months_vector. Print out factor_months_vector and assert that R prints out the factor levels below the actual values.

```{r}
months_vector <- c("March", "April","January", "November", "January", "September", "October", "September", "November", "August", "January", "November", "November", "February", "May", "August", "July", "December", "August", "August", "September", "November", "February", "April")
months_vector
```
```{r}
factor_months_vector <- factor(months_vector)
factor_months_vector
```
3. Then check the summary() of the months_vector and factor_months_vector. | Interpret the results of both vectors. Are they both equally useful in this case?
```{r}
summary(months_vector)
```
```{r}
summary(factor_months_vector)
```
```
The results of the summary of months_vector does not really show any contributtions. However, factor_months_vector gives a summary on how many number of times month is entered in the vector, which is handy when summarizing harge amounts of data.
```

4. Create a vector and factor for the table below.

```{r}
direction <- c("East", "West","North")
frequency <- c(1,4,3)
```
```{r}
direction_factor <- factor(direction, levels = c("East", "West", "North"))
direction_factor
```
```{r}
freq <- factor(frequency, levels = c("1", "4", "3"))
freq
``` 
5. Enter the data below in Excel with file name = import_march.csv
```{r}
importmarch <- read.csv(file="import_march.csv")
importmarch
```

a. Import the excel file into the Environment Pane using read.table() function.
Write the code.

```{r}
read_importmarch <- read.table(file ='/cloud/project/import_march.csv', header = TRUE, sep = ',')
read_importmarch
```
6. Full Search
```{r}
input <-  as.numeric(readline(prompt = "Enter a number from 1-50:"))

input <- 49
if(input == 20){
  print("TRUE")
} else if (input > 50){
  print("The number selected is beyond the range of 1 to 50.")
} else{
  print(input)
}
```

7. Change

```{r}
minBills <- function(price) {
  
#user input
price<- 100
  
  if (price %% 50 != 0) {
    return("The price must be a multiple of 50.")
  }
  
  bills <- c(500, 100, 50, 200, 1000)
  
  numBills <- 0
  
  for (i in 1:length(bills)) {
    while (price >= bills[i]) {
      price <- price - bills[i]
      numBills <- numBills + 1
    }
  }
  return(numBills)
}


price <- as.integer(readline(prompt="Enter the price of the snack (in pesos, divisible by 50): "))

print(paste("The minimum number of bills needed to pay for a", price, "peso snack is", minBills(price), "."))

```

8. The following is each student’s math score for one semester. Based on this, answer the following questions.

a. Create a dataframe from the above table. Write the R codes and its output.

```{r}
students_df <- data.frame(Name = c("Annie", "Thea", "Steve", "Hanna"),
                          Grade1 = c(85, 65, 65, 75),
                          Grade2 = c(65, 75, 95, 90),
                          Grade3 = c(65, 65, 100, 75),
                          Grade4 = c(75, 55, 90, 80))
students_df
```

b. Without using the rowMean function, output the average score of students whose average math score over 90 points during the semester. write R code and its output.

```{r}
average_scores <- (students_df$Grade1 + students_df$Grade2 + students_df$Grade3 + students_df$Grade4) / 4
good_students <- which(average_scores > 90)

for (i in 1:nrow(students_df)) {
 if (i %in% good_students) {
    cat(paste0(students_df$Name[i], "'s score this semester is "), average_scores[i], "\n")
 } else {
    cat(paste0(students_df$Name[i], "'s does not have a score of 90 and above."), "\n")
 }
}
```

c. Without using the mean function, output as follows for the tests in which the average score was less than 80 out of 4 tests. Example output: The nth test was difficult.

```{r}
count <- 0

for (i in 1:nrow(students_df)) {
  
  average_scores <- sum(students_df[i, -1]) / 4
  if (average_scores < 80) {
    count <- count + 1
    print(paste("The", count, "nth test was difficult for", students_df[i, 1], "."))
  }
}
```

d. Without using the max function, output as follows for students whose highest score for a semester exceeds 90 points. Example Output: Annie’s highest grade this semester is 95.
```{r}
highest_grade <- apply(students_df[, -1], 1, max)

students_above_90 <- students_df[highest_grade > 90, ]

for (i in 1:nrow(students_above_90)) {
  cat(students_above_90[i, 1], "'s highest grade this semester is", max(students_above_90[i, -1]), ".")
}
```


